---
title: "workshop-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## welcome to workshop-08
Today we will be covering the following topics:
- Github and Git
- Intro to tidyverse
  - tidy evaluation and metaprogramming
  - data frames and data masking, and why I didn't teach tidyverse first when we learned R
  - dplyr
  - brief discussion about other tidyverse libraries that might be of use (stringr, tidyr, plyr)
  - ggplot2 (if time permits)

## let's load the libraries  
```{r}
library(dplyr)
```

  
## the data we will be working with
```{r}
# courtesy of @gregory-golden;
# noised and randomised data from flow experiments; hw8_data.csv
df <- read.csv("hw8_data.csv", check.names = FALSE)

head(df)
colnames(df)
```

```{r}
smallDf <- select(df, donor, Sex, `Disease Status`)

filter(smallDf, Sex == "F")

mutate(smallDf, donorExpanded = paste0("SCUBA_", donor))

mutate(smallDf, Sex = ifelse(Sex == "M", "Male", Sex))

arrange(smallDf, desc(donor))
```

```{r}
jakeDf <- smallDf %>%
  filter(Sex == "F") %>%
  mutate(jakeWrath = 42) %>%
  select(-jakeWrath) %>%
  arrange(desc(donor))

```

```{r}
df %>%
  group_by(`Disease Status`, Tissue) %>%
  summarize(CD45posAvg = mean(`CD45+`))
```

